Camellia:

### 2012-10-08 docs/Nate/Proposal/presentation/tst_2012_talk.tex
Added TST 2012 talk.

\date{August 30, 2012}
\author[Nathan V. Roberts]{Nathan V. Roberts \\
Supervisors: Leszek Demkowicz, Robert Moser \\
Committee: Todd Arbogast, George Biros, Thomas Hughes, Venkatramanan Raman
}
\institute{Institute for Computational and Engineering Sciences\\
The University of Texas at Austin}
\title[DPG for Incompressible Flow]{A Discontinuous Petrov-Galerkin Methodology\\ for Incompressible Flow}
\subtitle{Towards Automatic, Robust Mesh Adaptivity}

\begin{document}
\begin{frame}
%\begin{center}
%\includegraphics[width=.8\linewidth]{grand_logo}\\
%\end{center}
\titlepage
%\begin{flushright}
%\includegraphics[scale=0.1]{asc_logo}\\
%\end{flushright}
\end{frame}

%===============================================================================
% ACKNOWLEDGEMENTS
%===============================================================================
%\begin{frame}
%\frametitle{Acknowledgments}
%
%Collaborators:
%\begin{itemize}
%\item Leszek Demkowicz (UT/PECOS)
%\item Denis Ridzal (Sandia/CSRI)
%\item Pavel Bochev (Sandia/CSRI)
%\item Jesse Chan (UT/PECOS)
%\end{itemize}
%
%Support:
%\begin{itemize}
%\item Sandia (summer internships)
%\item PECOS (GRA funding)
%\item ICES (CSEM Fellowship)
%\end{itemize}
%
%\end{frame}

\begin{frame}
\frametitle{Motivation}

Incompressible flows:
\begin{itemize}
\item arise in a variety of applications, from hydraulics to aerodynamics
\item Navier-Stokes equations are of fundamental physical and mathematical interest:

\begin{itemize}
\item believed to hold the key to understanding turbulence
\item precise conditions for existence and uniqueness of solutions remain unknown (Millennium Prize problem)
\end{itemize}

\end{itemize}
%\hspace{1cm} 
\begin{center}
\includegraphics[width=6cm]{../figures/submarine}
\end{center}
% \hfill \includegraphics{../figures/propeller}\hspace{1cm}
\end{frame}

\begin{frame}
\frametitle{Motivation}

\begin{center}\nocite{Williamson1996}
\includegraphics[scale=.5]{../figures/cylinderWakeRe300Williamson.png}\\
\end{center}

Typical solutions of incompressible flow problems involve both fine- and large-scale phenomena.

\begin{center}
$\implies$
\end{center}

A \pecosbold{uniform} finite element mesh of sufficient granularity is
\begin{itemize}
\item \pecosbold{wasteful} of computational resources, or
\item \pecosbold{infeasible} because of resource limitations.
\end{itemize}

\vspace{3mm}
\begin{center}
Therefore, an \pecosbold{adaptive mesh} is required.
\end{center}

@@

\begin{frame}
\frametitle{Motivation}
In industry:
\begin{itemize}
\item adaptivity schemes used are ad hoc, requiring a domain expert to predict features of the solution,
\item a badly chosen mesh may take considerably longer to converge or fail to converge, and
\item typically, the Navier-Stokes solve will be just one component in an optimization loop $\implies$ any failure requiring human intervention is costly.
\end{itemize}

\begin{center}
\includegraphics[scale=0.35]{../figures/Jesse/mesh10.png}
\end{center}

@@

\begin{frame}
\vspace{5mm}
\begin{center}
\pecosbold{Goal}: Develop a solver for the incompressible Navier-Stokes equations that provides \pecosbold{robust adaptivity} starting from a coarse mesh.
\end{center}
\pause
By \pecosbold{robust}, we mean:
\begin{itemize}
\item always \pecosbold{converges} to a solution in predictable time, and
\pause
\item the adaptive scheme is \pecosbold{independent} of the problem: no special expertise required for adaptivity.
\end{itemize}

% Explain here what we mean by robust: solver always converges to a solution in predictable time, and that the adaptive scheme is independent of the problem: no special expertise is required for adaptivity.

@@

%===============================================================================
% OUTLINE
%===============================================================================

% UNCOMMENT THIS TO ADD AN OUTLINE AT THE TOP OF EACH SECTION...
%\AtBeginSection[]
%{
%  \begin{frame}<beamer>
%    \frametitle{Outline}
%    \tableofcontents[currentsection]
%  \end{frame}
%}

\begin{frame}
\frametitle{Outline}
\tableofcontents
\end{frame}

\section{DPG in Brief} % 5-6 slides
% motivate as a min-residual method
% end with the two flow chart slides
%===============================================================================
% The Abstract Problem and Minimization of the Residual
%===============================================================================
\begin{frame}
\frametitle{The Abstract Problem and Minimization of the Residual}
Take $U,V$ Hilbert.\\
\vspace{5mm}
We seek $u \in U$ such that
\[
b(u,v) = l(v) \quad \forall v \in V,
\]
where $b$ and $l$ are linear in $v$.  Define $B$ by $Bu = b(u,\cdot) \in V'$; $Bu$ is a linear functional on the test space $V$.\\
\vspace{5mm}
We seek to minimize the residual in the discrete space $U_{h} \subset U$:
\begin{align*}
u_{h} = \underset{w_{h} \in U_{h}} \argmin \,\, \frac{1}{2} \norm{Bw_{h}-l}_{V'}^{2}.
\end{align*}
\end{frame}

\begin{frame}
\frametitle{The Abstract Problem and Minimization of the Residual}
\begin{align*}
u_{h} = \underset{w_{h} \in U_{h}} \argmin \,\, \frac{1}{2} \norm{Bw_{h}-l}_{V'}^{2}.
\end{align*}

% PATTER: ``not especially easy to work with'': non-constructive; involves taking a supremum
Now, the dual space $V'$ is not especially easy to work with; we would prefer to work with $V$ itself.  Recalling that the Riesz operator $R_{V} : V \rightarrow V'$ defined by 
\begin{align*}
\langle R_{V}v, \delta v\rangle=(v,\delta v)_{V}, \quad \forall \delta v \in V,
\end{align*}
is an \emph{isometry}---$\norm{R_{V}v}_{V'} = \norm{v}_{V}$---we can rewrite the term we want
to minimize as a norm in $V$:

\begin{align*}
\frac{1}{2} \norm{Bw_{h}-l}_{V'}^{2} &= \frac{1}{2} \norm{R_{V}^{-1}\left(Bw_{h}-l\right)}_{V}^{2}\\
                                                              &= \frac{1}{2} \left(R_{V}^{-1}\left(Bw_{h}-l\right), R_{V}^{-1}\left(Bw_{h}-l\right) \right)_{V}.
\end{align*}
\end{frame}

\begin{frame}
\frametitle{The Abstract Problem and Minimization of the Residual}
We seek to minimize
\begin{align*}
\frac{1}{2} \left(R_{V}^{-1}\left(Bw_{h}-l\right), R_{V}^{-1}\left(Bw_{h}-l\right) \right)_{V}.
\end{align*}
The first-order optimality condition requires that the G\^ateaux derivative be equal to zero for minimizer $u_{h}$; assuming $B$ is linear, we have
\begin{align*}
\left(R_{V}^{-1}\left(Bu_{h}-l\right), R_{V}^{-1}B \delta u_{h} \right)_{V} = 0, \quad \forall \delta u_{h} \in U_{h}.
\end{align*}
By the definition of $R_{V}$, this is equivalent to
\begin{align*} % PATTER: use the phrase ``duality pairing''
\langle Bu_{h} - l, R_{V}^{-1}B \delta u_{h} \rangle = 0 \quad \forall \delta u_{h} \in U_{h}.
\end{align*}

\begin{frame}
\frametitle{The Abstract Problem and Minimization of the Residual}
We have:
\begin{align*}
\langle Bu_{h} - l, R_{V}^{-1}B \delta u_{h} \rangle = 0 \quad \forall \delta u_{h} \in U_{h}.
\end{align*}
Now, if we identify $v_{\delta u_{h}} = R_{V}^{-1}B \delta u_{h}$ as a test function, we can rewrite this as
\begin{align*}
b(u_{h},v_{\delta u_{h}}) &= l(v_{\delta u_{h}}).
\end{align*}

Thus, the discrete solution that minimizes the residual is exactly attained by testing the original equation with appropriate test functions.  We call these \pecosbold{optimal test functions}.\footnote{\FootSize \bibentry{DPG2}}

\begin{frame}
\frametitle{Evaluating Error}
The derivation prompts the definition of an \pecosbold{energy norm} on the trial space:
\[
\NVRnorm{u}_{E} \NVReqdef \NVRnorm{Bu}_{V'} = \sup_{v \in V} \frac{b(u,v)}{\NVRnorm{v}_{V}}.
\]
We can use this to define a measure of the error:
\begin{align*}
\NVRnorm{u - u_{h}}_{E} &= \NVRnorm{Bu - Bu_{h}}_{V'} \\
&=  \NVRnorm{l - Bu_{h}}_{V'} \\
&= \NVRnorm{R_{V}^{-1}\left(l - Bu_{h}\right)}_{V}.
\end{align*}
Note that all the terms in the final expression are known, so that we can \pecosbold{evaluate}, not merely estimate, the error.  This drives adaptivity.
\end{frame}

\begin{frame}
\frametitle{Graph Test Norm}

For a strong operator $A$ with formal adjoint $A^{*}$, the \pecosbold{adjoint graph space} is
\[
H_{A^{*}} = \left\{ v \in L^{2}(\Omega) : A^{*}v \in L^{2}(\Omega) \right\}
\]

and the \pecosbold{(adjoint) graph norm} on the test space $V$ is given by
\[
\NVRnorm{v}_{\rm graph} = \NVRnorm{v}_{H_{A^{*}}} = \left( \NVRnorm{v}^{2} + \NVRnorm{A^{*}v}^{2}\right)^{1/2}.
\]

\vspace{5mm}
\begin{center}
E.g. if $A^{*} = \NVRgrad$, then $H_{A^{*}} = H^{1}$, and $\NVRnorm{v}_{H_{A^{*}}} = \NVRnorm{v}_{H^{1}}$.
\end{center}
\end{frame}

\begin{frame}
\frametitle{Key Result: Well-posedness $\implies$ Optimal Convergence}
Under modest technical assumptions (true for Stokes), we have\footnote{\FootSize \bibentry{DPGStokes}}\\
\[
\NVRnorm{Au} \geq \gamma \NVRnorm{u} \implies \sup_{v \in H_{A^{*}}} \frac{b((u,\widehat{u}),v)}{\NVRnorm{v}_{H_{A^{*}}}} \geq \gamma_{\rm DPG} \left(\NVRnorm{u}^{2} + \NVRnorm{\widehat{u}}^{2}_{\widehat{H}_{A}(\Gamma_{h})}\right)^{1/2}
\]
where $\gamma_{\rm DPG} = O(\gamma)$ is a mesh-independent constant, and $\NVRnorm{\cdot}_{\widehat{H}_{A}(\Gamma_{h})}$ is the minimum energy extension norm.\\
\pause
\vspace{5mm}
By Babu\v{s}ka's Theorem,
\[
\implies \left(\NVRnorm{u - u_{h}}^{2} + \NVRnorm{\widehat{u} - \widehat{u}_{h}}^{2}_{\widehat{H}_{A}(\Gamma_{h})}\right)^{1/2} \leq \frac{M}{\gamma_{\rm DPG}} \left( {\rm B.A.E.} \right).
\]
\pause
Suffices to show that $\NVRnorm{Au} \geq \gamma \NVRnorm{u}$ to prove \pecosbold{optimal} convergence rate!
\end{frame}

\section{Equations for Incompressible Flow}
\begin{frame}
\frametitle{Incompressible Flow Equations}
\begin{center}
\includegraphics[width=0.9\linewidth]{../figures/incFlowEqns}\\
\end{center}
\end{frame}

\section{Stokes} % 10-12 slides
% lit. review (including various classical elements, discussion of locking and non-convergence)
% Stokes analysis, summary results
% Numerical experiments: manufactured solution, lid-driven cavity flow problem
% Probably worth mentioning the naive norm results: drives home why the analysis is valuable, and the fact that the optimal test function lives in a very specific space.
%===============================================================================
% Stokes: Classical Problem
%===============================================================================
\begin{frame}
\frametitle{Classical Stokes Problem}
The classical strong form of the Stokes problem in $\Omega \subset
\mathbb{R}^{2}$ is given by
\begin{align*}
- \mu \Delta \vect{u} + \NVRgrad p &= \vect{f} & \text{ in } \Omega, \\
\NVRdiv \vect{u} &= 0 & \text{ in } \Omega, \\
\vect{u} &= \vect{u}_D & \text{ on } \partial\Omega,
\end{align*}
where $\mu$ is viscosity, $p$ pressure, $\vect{u}$ velocity, and
$\vect{f}$ a vector forcing function.  Since by appropriate
non-dimensionalization we can eliminate the constant $\mu$, we take
$\mu=1$ throughout. 
\end{frame}
%===============================================================================
% Stokes: ``Literature Review:'' What's generally done, etc.
%===============================================================================
\begin{frame}
\frametitle{Stokes: Existing Approaches}

Naive discretizations of Stokes can exhibit \pecosbold{non-convergence} or \pecosbold{locking}.\footnote{\FootSize \bibentry{BoffiBrezziFortin}}
\\
\vspace{5mm}
Standard Galerkin discretizations:
\begin{itemize}
\item need to satisfy the \pecosbold{LBB condition}
\item examples: the MINI element, Crouzeix-Raviart element, the class of $Q_{k}-P_{k-1}$ elements
\item polynomial degree in the pressure discretization one lower than that for velocity $\implies$ pressure converges more slowly
\end{itemize}
%\item When solving saddle-point systems using Bubnov-Galerkin methods, there are two \pecosbold{Brezzi conditions} which must be satisfied.  The challenging one for Stokes is the \pecosbold{LBB condition}---much of the art of solving Stokes is to find discrete spaces that satisfy the LBB condition.%\footnote{\bibentry{BoffiBrezziFortin}}

%\item Some choices that satisfy the LBB condition are the MINI element, Crouzeix-Raviart element, and the class of $Q_{k}-P_{k-1}$ elements.  All of these require a polynomial degree in the pressure discretization one lower than that for velocity (so that pressure converges more slowly).

\begin{frame}
\frametitle{Stokes: Existing Approaches}
Local discontinuous Galerkin (LDG) method:\footnote{\FootSize \bibentry{CockburnKanschatSchotzauSchwab03}}
\begin{itemize}
\item \pecosbold{locally conservative}
\item allows pressure and velocity spaces to be chosen independently
\item equal-order spaces may be used
\item convergence \emph{rate} for the pressure remains lower than that for the velocity
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Stokes: Existing Approaches}
Divergence-conforming B-splines:\footnote{\FootSize \bibentry{EvansThesis}}
\begin{itemize}
\item pointwise divergence-free space $\implies$ locally conservative
\item equal-order spaces
\item optimal convergence rates for pressure \pecosbold{and} velocity
\item non-conforming for multi-patch domains $\implies$ DG techniques for tangential continuity across patch interfaces
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{DPG Applied to Stokes}

To apply DPG, we need a first-order system.  We introduce $\NVRtensor{\sigma}= \NVRgrad
\vect{u}$:
\begin{align*}
- \NVRdiv \NVRtensor{\sigma} + \NVRgrad p &= \vect{f} & \text{ in } \, \Omega, \\
 \NVRdiv \vect{u} &= 0 & \text{ in } \, \Omega, \\
 \NVRtensor{\sigma} - \NVRgrad \vect{u} &= 0 & \text{ in } \, \Omega.
\end{align*}
Testing with $(\vect{v},q,\vect{\tau})$, and integrating by parts, we have
\begin{align*}
\left(\NVRtensor{\sigma} - p \NVRtensor{I}, \NVRgrad \vect{v} \right)_{\Omega_{h}} - \left\langle \widehat{\vect{t}}_{n}, \vect{v} \right\rangle_{\Gamma_{h}} &= (\vect{f}, \vect{v} )_{\Omega_{h}}\\
\left(\vect{u}, \NVRgrad q \right)_{\Omega_{h}} -  \left\langle \widehat{\vect{u}} \cdot \vect{n}, q  \right\rangle_{\Gamma_{h}} &= 0\\
\left( \NVRtensor{\sigma}, \NVRtensor{\tau} \right)_{\Omega_{h}} + \left( \vect{u}, \NVRdiv \NVRtensor{\tau} \right)_{\Omega_{h}} - \left\langle \widehat{\vect{u}}, \NVRtensor{\tau} \vect{n}  \right\rangle_{\Gamma_{h}} &= \NVRtensor{0},
\end{align*}
where traction $\vect{t}_{n} \eqdef ( \NVRtensor{\sigma} - p \NVRtensor{I} ) \vect{n}$, and the hatted variables $\widehat{\vect{t}}_{n}$ and $\widehat{\vect{u}}$ are new unknowns representing the traces of the corresponding variables at the boundary.

\begin{frame}
\frametitle{DPG Applied to Stokes}
DPG Formulation:
\begin{align*}
b(u,v) =& \left(\NVRtensor{\sigma} - p \NVRtensor{I}, \NVRgrad \vect{v} \right)_{\Omega_{h}} - \left\langle \widehat{\vect{t}}_{n}, \vect{v} \right\rangle_{\Gamma_{h}}\\
&+\left(\vect{u}, \NVRgrad q \right)_{\Omega_{h}} -  \left\langle \widehat{\vect{u}} \cdot \vect{n}, q  \right\rangle_{\Gamma_{h}}\\
&+ \left( \NVRtensor{\sigma}, \NVRtensor{\tau} \right)_{\Omega_{h}} + \left( \vect{u}, \NVRdiv \NVRtensor{\tau} \right)_{\Omega_{h}} - \left\langle \widehat{\vect{u}}, \NVRtensor{\tau} \vect{n}  \right\rangle_{\Gamma_{h}} = (\vect{f}, \vect{v} )_{\Omega_{h}} = l(v).
\end{align*}
The natural spaces for the trial variables are then:
\begin{itemize}
\item fields: $p \in L^{2}(\Omega), \vect{u} \in \vect{L}^{2}(\Omega), \NVRtensor{\sigma} \in \vect{L}^{2}(\Omega)$,
\item fluxes: $\widehat{\vect{t}}_{n} \in \NVRtensor{H}^{-1/2}(\Gamma_{h})$,
\item traces: $\widehat{\vect{u}} \in \vect{H}^{1/2}(\Gamma_{h})$.
\end{itemize}

The natural norms for fluxes and traces are \emph{minimum energy extension norms}.

\begin{frame}
\frametitle{Polynomial Space Choices}
How to select polynomial spaces for $u$ and $\widehat{u}$?
\begin{itemize}
\item Graph norm $\implies$ error bounded by best approximation error.
\vspace{3mm} \pause
\item Want all variables to have same B.A.E. convergence rates.
\vspace{3mm} \pause
\item Fix polynomial order $k$ for $L^{2}$ variables.  ($k+1$ rate.)
\vspace{3mm} \pause
\item $\implies$ $k+1$ for $H^{1/2}(\Gamma_{h})$.
\vspace{3mm} \pause
\item $\implies$ $k$ for $H^{-1/2}(\Gamma_{h})$.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Graph Test Norm}
The adjoint graph norm for our Stokes formulation is:\footnote{\FootSize \bibentry{DPGStokes}}
\begin{align*}
\norm{(\NVRtensor{\tau},\vect{v}, q)}_{\rm graph}^{2} =& \norm{\NVRdiv \NVRtensor{\tau} - \NVRgrad q}^{2} + \norm{\NVRdiv \vect{v}}^{2} + \norm{ \NVRtensor{\tau} + \NVRgrad{\vect{v}}}^{2}\\ 
&+ \norm{\NVRtensor{\tau}}^{2} + \norm{\vect{v}}^{2} + \norm{q}^{2}.
\end{align*}

\begin{frame}
\frametitle{Manufactured Solution}
For our first numerical experiment (and following Cockburn et al.\footnote{\FootSize \bibentry{CockburnKanschatSchotzauSchwab03}}), we consider a manufactured solution
\begin{align*}
u_{1} &=  -e^{x} ( y \cos y + \sin y )\\
u_{2} &=  e^{x}  y \sin y\\
p &= 2 \mu e^{x} \sin y
\end{align*}
on domain $\Omega=(-1,1)^{2}$.  We use this to determine appropriate boundary conditions for the DPG problem.  We also perform an $L^{2}$ projection of the exact solution into the trial space to find the solution with the best approximation error.
\end{frame}

\begin{frame}
\frametitle{Graph Test Norm: $u_{1}$ convergence}
\begin{figure}[!htb]
\center
{\setlength{\fboxsep}{1pt}\colorbox{pecos2}{\includegraphics[scale=.38]{../figures/u1_graph_h.png}}}
\caption{$h$-convergence with the graph norm: $u_{1}$.  Dashed lines: best approximation error.}
\end{figure}
\end{frame}

\addtocounter{framenumber}{-1} % make the next frame # the same as the previous
\begin{frame}
\frametitle{Graph Test Norm: $u_{2}$ convergence}
\begin{figure}[!htb]
\center
{\setlength{\fboxsep}{1pt}\colorbox{pecos2}{\includegraphics[scale=.38]{../figures/u2_graph_h.png}}}
\caption{$h$-convergence with the graph norm: $u_{2}$.  Dashed lines: best approximation error.}
\end{figure}
\end{frame}

\addtocounter{framenumber}{-1} % make the next frame # the same as the previous
\begin{frame}
\frametitle{Graph Test Norm: $p$ convergence}
\begin{figure}[!htb]
\center
{\setlength{\fboxsep}{1pt}\colorbox{pecos2}{\includegraphics[scale=.38]{../figures/pressure_graph_h.png}}}
\caption{$h$-convergence with the graph norm: $p$.  Dashed lines: best approximation error.}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Naive Test Norm\footnote{\FootSize \bibentry{RobertsetAl10}}}
What if we don't use the graph norm, but a naive choice instead?

\begin{align*}
\norm{(\NVRtensor{\tau},\vect{v}, q)}_{\rm naive}^{2} = \norm{\NVRtensor{\tau}}^{2} + \norm{\NVRdiv \NVRtensor{\tau}}^{2} + \norm{\vect{v}}^{2} + \norm{\NVRgrad \vect{v}}^{2} + \norm{q}^{2} + \norm{\NVRgrad q}^{2}.
\end{align*}
\end{frame}

\begin{frame}
\frametitle{Naive Test Norm: $u_{1}$ convergence}
\begin{figure}[!htb]
\center
{\setlength{\fboxsep}{1pt}\colorbox{pecos2}{\includegraphics[scale=.38]{../figures/u1_naive_h.png}}}
\caption{$h$-convergence with the naive norm: $u_{1}$.  Dashed lines: best approximation error.}
\end{figure}
\end{frame}

\addtocounter{framenumber}{-1} % make the next frame # the same as the previous
\begin{frame}
\frametitle{Naive Test Norm: $u_{2}$ convergence}
\begin{figure}[!htb]
\center
{\setlength{\fboxsep}{1pt}\colorbox{pecos2}{\includegraphics[scale=.38]{../figures/u2_naive_h.png}}}
\caption{$h$-convergence with the naive norm: $u_{2}$.  Dashed lines: best approximation error.}
\end{figure}
\end{frame}

\addtocounter{framenumber}{-1} % make the next frame # the same as the previous
\begin{frame}
\frametitle{Naive Test Norm: $p$ convergence}
\begin{figure}[!htb]
\center
{\setlength{\fboxsep}{1pt}\colorbox{pecos2}{\includegraphics[scale=.38]{../figures/pressure_naive_h.png}}}
\caption{$h$-convergence with the naive norm: $p$.  Dashed lines: best approximation error.}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Graph vs. Naive Test Norm}
What's the difference between the two norms?  Why are the results better with the graph norm?

\begin{align*}
\norm{(\NVRtensor{\tau},\vect{v}, q)}_{\rm naive}^{2} =& \norm{\NVRdiv \NVRtensor{\tau}}^{2} + \norm{\NVRdiv \vect{v}}^{2} + \norm{\NVRgrad q}^{2} + \norm{\NVRtensor{\tau}}^{2} + \norm{\vect{v}}^{2} + \norm{q}^{2}\\
\vspace{30mm}
\norm{(\NVRtensor{\tau},\vect{v}, q)}_{\rm graph}^{2} =& \norm{\NVRdiv \NVRtensor{\tau} - \NVRgrad q}^{2} + \norm{\NVRdiv \vect{v}}^{2} + \norm{ \NVRtensor{\tau} + \NVRgrad{\vect{v}}}^{2}\\ 
&+ \norm{\NVRtensor{\tau}}^{2} + \norm{\vect{v}}^{2} + \norm{q}^{2}
\end{align*}
\pause
The naive norm is \pecosbold{stronger}---e.g. it requires $\NVRdiv \NVRtensor{\tau} \in L^{2}$ and $\NVRgrad q \in L^{2}$, whereas the graph norm merely requires that $\NVRdiv \NVRtensor{\tau} - \NVRgrad q \in L^{2}$.

\begin{frame}
\frametitle{Lid-Driven Cavity Flow Problem}
A classic test case for Stokes flow is the \pecosbold{lid-driven cavity flow problem}.  Consider a square cavity with an incompressible, viscous fluid, with a lid that moves at a constant rate.  The resulting flow will be vorticular; there will be \pecosbold{Moffat eddies} at the corners; in fact, the exact solution will have an infinite number of such eddies, visible at progressively finer scales.\footnote{\FootSize \bibentry{Moffat}}
\begin{center}
{\includegraphics[scale=0.40]{../figures/cavity_flow_cartoon.pdf}}
\end{center}
\end{frame}

\begin{frame}
\frametitle{Lid-Driven Cavity Flow Problem}
\begin{center}
{\includegraphics[scale=0.40]{../figures/cavity_flow_cartoon.pdf}}
\end{center}
\small{Because the BCs for the problem are discontinuous at the top corners, the exact solution lies outside of $H^{1}$.  So we introduce a small ``ramp'' on either side, of width $\epsilon = \frac{1}{64}$.}
\end{frame}

\begin{frame}
\frametitle{Lid-Driven Cavity Flow Problem}
Since the exact solution is unknown, we compare an overkill mesh to a series of adaptive and uniformly refined solutions.\\

\begin{itemize}
\item For all meshes, we use quadratic field variables ($k=2$).
\item Overkill mesh had $256 \times 256$ elements (5,576,706 dofs).
\item Initial adaptive mesh had $2 \times 2$ elements.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Lid-Driven Cavity Flow: $h$-adaptivity}
\begin{figure}[!htb]
\begin{center}
{\setlength{\fboxsep}{1pt}\colorbox{pecos2}{\includegraphics[scale=0.40]{../figures/adaptive_cavity_flow_quadratic_vs_overkill.pdf}}}
\end{center}
\caption{Euclidean norm of $L^{2}$ error in all field variables in $h$-adaptive mesh relative to an overkill mesh with $256 \times 256$ quadratic elements.  The Euclidean norm of the overkill solution is 6.73.}
\end{figure}

\begin{frame}
\frametitle{Lid-Driven Cavity Flow: Streamlines}
\begin{center}
{\setlength{\fboxsep}{1pt}\colorbox{pecos2}{\includegraphics[scale=0.25]{../figures/streamlines_p2_r7.pdf}}} \hspace{5mm} {\setlength{\fboxsep}{1pt}\colorbox{pecos2}{\includegraphics[scale=0.25]{../figures/streamlines_detail_p2_r7.pdf}}}
\end{center}
Streamlines for the full cavity and for the lower-left corner, on a quadratic mesh after 7 adaptive refinements.  The lower-left corner shows the first Moffat eddy.  The final mesh has 124 elements and 11,202 dofs.
\end{frame}

>>>>

\date{TST Meeting\\ October 17, 2012}
\author[Nate Roberts]{Nate Roberts \\
Collaborators: Leszek Demkowicz, Jesse Chan, Truman Ellis
}
\institute{Institute for Computational and Engineering Sciences\\
The University of Texas at Austin}
\title[Camellia: A DPG Toolbox]{Camellia:}
\subtitle{A Toolbox for a Class of Discontinuous Petrov-Galerkin Methods\\ Using Trilinos}
\begin{document}
\begin{frame}
%\begin{center}
%\includegraphics[width=.8\linewidth]{grand_logo}\\
%\end{center}
\titlepage
%\begin{flushright}
%\includegraphics[scale=0.1]{asc_logo}\\
%\end{flushright}
\end{frame}

\section{DPG in Brief} % 3-4 slides
% motivate as a min-residual method
% end with the two flow chart slides
%===============================================================================
% The Abstract Problem and Minimization of the Residual
%===============================================================================
\begin{frame}
\frametitle{The Abstract Problem and Minimization of the Residual}\small
Take $U,V$ Hilbert.  We seek $u \in U$ such that
\[
b(u,v) = l(v) \quad \forall v \in V,
\]
where $b$ is bilinear and $l$ is linear in $v$.  Define $B$ by $Bu = b(u,\cdot) \in V'$.\\
\vspace{5mm}
We seek to minimize the residual in the discrete space $U_{h} \subset U$:
\begin{align*}
u_{h} = \underset{w_{h} \in U_{h}} \argmin \,\, \frac{1}{2} \norm{Bw_{h}-l}_{V'}^{2}.
\end{align*}

If we use test functions $v_{\delta u_{h}} = R_{V}^{-1}B \delta u_{h}$, then the discrete solution to
\begin{align*}
b(u_{h},v_{\delta u_{h}}) &= l(v_{\delta u_{h}}) \forall \delta u_{h} \in U_{h}
\end{align*}
minimizes the residual.  We call the $v_{\delta u_{h}}$ \pecosbold{optimal test functions}.\footnote{\FootSize \bibentry{DPG2}}

@@

\begin{frame}
\frametitle{Key Features of DPG}\small
\begin{itemize}
\item \pecosbold{guarantees} minimum residual in a norm that depends on the test space, \pecosbold{a free choice}
\item for non-singularly-perturbed, well-posed linear problems, \pecosbold{provably optimal} convergence rates when the ``adjoint graph norm'' is used for the test space
\item \pecosbold{SPD} stiffness matrix
\item discontinuous test space $\implies$ solving for test functions is a \pecosbold{local} operation
\item nonlinear problems: so far, apply the linear method to the linearized problem
\item have also derived a method that instead minimizes the nonlinear residual
\end{itemize}
\begin{block}{Adjoint Graph Norm}
For a strong operator $A$ with formal adjoint $A^{*}$, the \pecosbold{adjoint graph norm} on the test space $V$ is given by
\[
\NVRnorm{v}_{\rm graph} = \NVRnorm{v}_{H_{A^{*}}} = \left( \NVRnorm{v}^{2} + \NVRnorm{A^{*}v}^{2}\right)^{1/2}.
\]
\end{block}
\end{frame}

\section{Camellia} % 10-12 slides
\begin{frame}[fragile]
\frametitle{Camellia\footnote{\FootSize \bibentry{RobertsetAl11}}}
\pecosbold{Design Goal}: make DPG research and experimentation as simple as possible, without sacrificing too much by way of performance.
\vspace{3mm}
\begin{center}
\includegraphics[scale=.65]{../figures/CamelliaVennDiagram.pdf}
\end{center}
\addtocounter{footnote}{1}
\footnotetext{\FootSize \bibentry{TrilinosShortAuthors}}
\end{frame}

\begin{frame}
\frametitle{Camellia: 2011 Design}
\begin{itemize}
\item Camellia defines \pecosbold{abstract classes} for bilinear form, inner product, right-hand side, boundary conditions.
\item User implements problem-specific code through \pecosbold{subclassing}.
\item Allows \pecosbold{automatic} definition of adjoint graph norm.
\end{itemize}

\begin{table}
\begin{center}
\begin{tabular}{| c | c |}
\hline
Implementation & Lines of Code\\
\hline
Stokes bilinear form & 341 \\
Stokes RHS and BCs & 251 \\
Automatic graph norm & 198 \\
\hline
\end{tabular}
\end{center}
\end{table}
\end{frame}

\begin{frame}
\frametitle{Camellia: 2012 Design}
\begin{itemize}
\item Key observation: bilinear form, inner product, and RHS can all be defined using \pecosbold{linear terms} involving the trial and test space variables.
\item We can define a linear term as \code{lt = f * v} where \code{f} is a function and \code{v} is a variable.
\item Functions can depend on spatial coordinates, and may be scalar-, vector-, or tensor-valued.
\item Variables might be operated on by a linear differential operator.
\item Using these as primitives, Camellia allows \pecosbold{declarative} specification of bilinear form, inner product, and RHS.
\end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{Camellia: 2012 Design, Poisson implementation}
Recall our Poisson formulation:
\begin{align*}
\text{``}b(u,v)\text{''} = (\widehat{\psi}_n,v)_{\Gamma_h} - (\vect{\psi}, \nabla v)_{\Omega_h} &\\
+ (\vect{\psi}, \vect{q})_{\Omega_h} + (\widehat{\phi}, \vect{q}  \cdot \vect{n})_{\Gamma_h} - (\phi, \nabla \cdot \vect{q})_{\Omega_h} &= (f,v)_{\Omega_h}.
\end{align*}
\begin{lstlisting}
  VarFactory varFactory; 
  VarPtr v = varFactory.testVar("v", HGRAD);
  VarPtr q = varFactory.testVar("q", HDIV);

  VarPtr psi_hat_n = varFactory.fluxVar("\\widehat{\\phi}_n");
  VarPtr phi_hat = varFactory.traceVar("\\widehat{\\psi}");
  VarPtr psi = varFactory.fieldVar("\\psi", VECTOR_L2);
  VarPtr phi = varFactory.fieldVar("\\phi", L2);
\end{lstlisting}

@@

\begin{frame}[fragile]
\frametitle{Camellia: 2012 Design, Poisson implementation}
Recall our Poisson formulation:
\begin{align*}
\text{``}b(u,v)\text{''} = (\widehat{\psi}_n,v)_{\Gamma_h} - (\vect{\psi}, \nabla v)_{\Omega_h} &\\
+ (\vect{\psi}, \vect{q})_{\Omega_h} + (\widehat{\phi}, \vect{q}  \cdot \vect{n})_{\Gamma_h} - (\phi, \nabla \cdot \vect{q})_{\Omega_h} &= (f,v)_{\Omega_h}.
\end{align*}
\begin{lstlisting}
  BFPtr bf = Teuchos::rcp( new BF(varFactory) );

  bf->addTerm(psi_hat_n, v);
  bf->addTerm(-psi, v->grad());

  bf->addTerm(psi, q);
  bf->addTerm(phi_hat, q->dot_normal());
  bf->addTerm(-phi, q->div());
\end{lstlisting}

RHS, BCs, and test space norms can be specified in a similar fashion.

@@

\begin{frame}
\frametitle{Camellia: 2012 vs. 2011 Design, Lines of Code}

\begin{table}[h!b!p!]
\begin{center}
\begin{tabular}{| c | c | c |}
\hline
Implementation & Lines of Code, Old & Lines of Code, New\\
\hline
Stokes bilinear form & 341 & 50 \\
Stokes RHS and BCs & 251 & 90\\
Automatic graph norm & 198 & 36 \\
\hline
\end{tabular}
\end{center}
\end{table}

\begin{block}{Camellia: Future Work}
\begin{itemize}
\item support for curved geometry.
\item minimize the \emph{nonlinear} residual (adds a hessian term).
\item in further future: 1D and 3D mesh support, distributed mesh and solution storage.
\end{itemize}
\end{block}
\end{frame}

%===============================================================================
% THANK YOU
%===============================================================================
\begin{frame}
\frametitle{}
\begin{block}{}
\center{Thank you!} \\
\center{Questions?}\\
\end{block}
\vspace{1cm}
\center{For more info:\\
nroberts@ices.utexas.edu}\\
\end{frame}
